---
title: "Regression tree"
author: "Li-Hsin Chien"
date: "2023-12-26"
output: html_document
---

## Chapter 16 Trees

### Section 16.1 Regression Trees

資料: *ozone* :A study the relationship between atmospheric ozone concentration and meteorology in the Los Angeles Basin in 1976. A number of cases with missing variables have been removed for simplicity.

讀入資料
```{r }
library(faraway)
data(ozone)
?ozone
```

Tree 分析套件: *rpart*

```{r}

#install.packages("rpart")
library(rpart)

(tmod<-rpart(O3~.,ozone))

```

第一個 split 為 temp 變數 (溫度)，切點為 67.5; 項目內容依序為 ("node 編號", "變數+切點", "node 內樣本數", "deviance = RSS","node 預測值")

*代表 terminal node

```{r}
plot(tmod)
text(tmod)
plot(tmod, compress=T, uniform=T, branch=.4)
text(tmod)
```

```{r}
#install.packages("rpart.plot")
library(rpart.plot)
prp(tmod,         # 模型
    faclen=0,           # 呈現的變數不要縮寫
    fallen.leaves=TRUE, # 讓樹枝以垂直方式呈現
    shadow.col="gray",  # 最下面的節點塗上陰影
    # number of correct classifications / number of observations in that node
    extra=1)  
```

```{r}
plot(jitter(predict(tmod)),residuals(tmod),xlab="Fitted",ylab="Residuals")
abline(h=0)

```

利用 Tree model 預測新樣本 (新樣本的covariate 為 median)

```{r}
(x0<- apply(ozone[,-1],2,median))
predict(tmod,data.frame(t(x0)))
```

### Section 16.2 Tree Pruning 

```{r}
set.seed(123)
tmode <-rpart(O3~., ozone, cp=0.001)
printcp(tmode)
plotcp(tmode)
```



```{r}
rpart.plot(tmod,type=3)
```

```{r}
tmodr<-prune.rpart(tmod,0.0154)
1-sum(residuals(tmodr)^2)/sum((ozone$O3-mean(ozone$O3))^2)

```


```{r}
set.seed(123)
tmod <- rpart(O3~., ozone[sample(330,165),])
(tmods <- prune.rpart(tmod,0.0154))
```

