---
title: "ELM_Ch2"
author: "Li-Hsin Chien"
date: "2023-11-27"
output: html_document
---


### Heart Disease Example

3154 healthy young men aged 39-59 from the San Francisco area were assessed for their personality type. All were free from coronary heart disease at the start of the research. Eight and a half years later change in this situation was recorded.



```{r }
library(faraway)

data(wcgs, package="faraway")

head(wcgs)


```

### 2.2 Fit logistic regression model

$y$: chd, 是否有冠狀動脈心臟病 (coronary heat disease)

$x=(height, cigs)$: 身高(height), 每天抽幾支菸 (cigs)

$$ log(odds)=log(\frac{p}{1-p})=\beta_0+\beta_1 x_1 +\beta_2 x_2 $$

```{r}
summary(wcgs$chd) #factor, levels: no yes, 會根據字母順序coding 成 0 & 1 (no & yes)
lmod <- glm(chd~height + cigs, family = binomial, wcgs)
summary(lmod)
(b<-lmod$coefficients)  #MLE beta_hat
(b.cov<-summary(lmod)$cov.unscaled) #MLE var(beta_hat)
```

$$ odds=e^{\beta_0}\cdot e^{\beta_1 x_1}\cdot e^{\beta_2 x_2} $$
```{r}
exp(b)
```

#### Q: 每天多抽一包菸 (20 支) 的影響?

```{r}
exp(b[3]*20)
```

#### A: 每天多抽一包菸，會使得得到心臟病的 odds 增加 59%。(會使odds變成原來的 1.59倍)

### 2.3 Inference

Deviance for model S, $D_S$
$$ D= 2 \cdot \{ \log L_L(\hat{\beta}_L)-\log L_S (\hat{\beta}_S)\}=-2\sum_{i=1}^n {y_i \cdot logit(\hat{p}_i) + log(1-\hat{p}_i)}  $$
where $L_L$ a saturated model, $\hat{p}_L=y_i$

#### likelihood ratio test

$H_0: \beta_1=0 (S)$ ; $H_1: o.w. (L)$, test statistic 
$$ 2\cdot log(\frac{L_L}{L_S})=D_S-D_L \sim \chi^2_{df=L\backslash S=1} $$

```{r}
summary(lmod)

(D_L <- lmod$deviance)

lmod.s<-glm(chd~ cigs, family = binomial, wcgs)
D_S <- lmod.s$deviance

t.lr<-D_S-D_L
1-pchisq(t.lr,1)

anova(lmod.s,lmod, test="Chi")

drop1(lmod,test="Chi")
```

#### Wald test

```{r}
b2<-b[2]
b2.sd<-sqrt(diag(b.cov)[2])
(t.wald<-b2/b2.sd) #H_0: beta_1=0
2*(1-pnorm(t.wald))

summary(lmod)
```

#### CI

$(1-\alpha)%$ CI for $\beta_i$: 
$$ \hat{\beta}_i \pm z^{\alpha/2} se(\hat{\beta}_i) $$
$(1-\alpha)%$ CI for $e^{\beta_i}$: 
$$ exp\{\hat{\beta}_i \pm z^{\alpha/2} se(\hat{\beta}_i)\} $$

```{r}
b2 + c(-1,1)*1.96*b2.sd
confint(lmod)

exp(b2 + c(-1,1)*1.96*b2.sd)
exp(confint(lmod))

```

### 2.4 diagnostics

```{r}
halfnorm(hatvalues(lmod))
wcgs[ which(hatvalues(lmod)>0.015),c("height","cigs","chd")]

par(mfrow=c(1,2))
halfnorm(hatvalues(lmod)[wcgs$chd=="yes"],main="chd=yes")
halfnorm(hatvalues(lmod)[wcgs$chd=="no"],main="chd=no")
```

### 2.5 model selection

$AIC=-2logL+2q=deviance+2q$

```{r}
wcgs$bmi <- with(wcgs, 703*wcgs$weight/(wcgs$height^2))
lmod <- glm(chd ~ age + height + weight +bmi + sdp + dbp + chol + dibep + cigs +arcus, family=binomial, wcgs)
lmodr <- step(lmod, trace=1) # trace=0: 不顯示過程
sumary(lmodr)

```

### 2.6 goodness of fit

#### calibration: $p_i$ 是否有估準?

```{r}
library(plyr)
wcgsm <- na.omit(wcgs) #去除有 missing 的資料
wcgsm <- mutate(wcgsm, predprob=predict(lmod,type="response")) #增加predprob變數

cut.point<-unique(quantile(wcgsm$predprob, (1:10)/11))

p.g<-cut(wcgsm$predprob, breaks=cut.point)
pred.p.g<-tapply(wcgsm$predprob,p.g,mean)
obs.n.g<-tapply(wcgsm$chd,p.g,function(x) sum(x=="yes"))
m.g<-tapply(wcgsm$chd,p.g,length)
obs.p.g<-obs.n.g/m.g
sd.g<-sqrt(pred.p.g*(1-pred.p.g)/m.g)

plot(pred.p.g,obs.p.g,xlab="Predicted prob.",ylab="Observed prob.",ylim=c(0,0.2))

apply(cbind(pred.p.g,obs.p.g,sd.g),1,function(x) segments(x[1],x[2]-2*x[3],x[1],x[2]+2*x[3],col="gray"))
abline(a=0,b=1,col=2)
```

#### discremination: $y_i$ 是否有分好?

```{r}
#install.packages("pROC")
library(pROC)
roc.lmod<-roc(data=wcgsm,response=chd,predictor=predprob)
auc(roc.lmod)
plot(roc.lmod)
```

