---
title: "hw1"
author: "Li-Hsin Chien"
date: "2024-01-16"
output: html_document
---

*The dataset teengamb concerns a study of teenage gambling in Britain. Fit a regression model with the expenditure on gambling as the response and the sex, status, income and verbal score as predictors. Present the output.*

```{r}
library(faraway)
data(teengamb)
r<-lm(gamble~sex+status+income+verbal, data=teengamb)
summary(r)

```

*(a) What percentage of variation in the response is explained by these predictors?*

$R^2 = 0.5267$. The percentage of variation in the response is 52.7%. 

*(b) Which observation has the largest (positive) residual? Give the case number.*

```{r}
r.res<-r$residuals
max(r.res)
which.max(r.res)
```

The observationcase with the case number 24 has the largest positive residual (94.25).

*(c) Computet the mean and median of the residuals. *

```{r}
mean(r.res)
median(r.res)
```

The mean of the residuals = 0.
The median of the residuals = -1.451.

*(d) Compute the correlation of the residuals with the fitted values.*

```{r}
cor(r$fitted.values, r.res)
```

The correlation of the residuals with the fitted values = 0.

*(e) Compute the correlation of the residuals with the income.*

```{r}
cor(teengamb$gamble, r.res)
```

The correlation of the residuals with the income = 0.688.

*(f) For all other predictors held constant, what would be the difference in predicted expenditure on gambling for a male comapred to a female?*

male: sex = 0

female: sex = 1

Therefore, for all other predictors held constant, the difference in predicted expenditure on gampling for a male compared to a female is gamle(sex=0) - gamle(sex=1) = 22.118.